{\plant}: A package for modelling forest trait ecology \& evolution

### 2015-08-28 ms.tex
Add lma numbers to the figure caption

While we have implemented a particular physiological model (\texttt{FFW16}), the
package is designed to allow arbitrary additional physiologies to be
introduced. Users are able to modify the physiological model in two
different ways. First, one can vary the parameters of the \texttt{FFW16} model
directly (see Table XXX for list, for examples see Appendix
\ref{sec:code}). Second, we allow for changes in key parameters to
bring about changes in other parameters (a hyper-parameterisation), enabling
straightforward modelling of trade-offs. For example, the trait \textsc{lma}
(leaf mass per unit leaf area) is used to estimate the rate of leaf turnover
($k_l$), based on an observed scaling relationship spanning across diverse
vegetation and plant functional types \citep{Wright-2004}:
$$k_l = \beta_{kl1} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{kl2}},$$
where $k_l$ and $\phi$ (lma) are both parameters of the core physiological
model, and $\beta_{kl1}, \beta_{kl2}, \phi_0$ are hyper-parameters. Such
linkages are defined within a user-supplied hyper-parameterisation function
(see Appendix \ref{sec:FFW16} for details). Finally, rather than varying the
parameters of the existing equations in the \texttt{FFW16} model, the
equations themselves can be changed and the higher level functionality of the
package can still be used.  Changed equations would amount to defining a new
physiological model in a new C++ file. As an example, in
the \texttt{FFW16b} physiology, we have modified the function determining
allocation to reproduction.

@@

The {\plant} model offers two methods for modelling the dynamics of \(N(x,
H, a)\) in a competing population. In the first ``stochastic'' mode, users
specify an average seed rain and patch size, and {\plant} then generates a
vector of seed arrival events and simulates stochastic
development of the resulting finite-sized population (see Appendices XXX for an
example).

@@

Increased interest in capturing the behaviour of size-structured  populations
has prompted development of several tools, each with its own distinct
strengths. The \textsc{EBT} software (see De Roos XXXX in this special feature) provides
a flexible framework that has been used to study a range of marine organisms.
Compared to the \textsc{EBT}, {\plant} is more focussed on plants, and
to that end offers refinements needed to handle modelling of the strong
size-asymmetric feedbacks that occur in plant communities, as well as the
extension to structured metapopulation. Other \textsc{TSPMs} also exist for modelling of vegetation, like \textsc{ED}
\citep[ver 1 and 2][]{Moorcroft-2001, Medvigy-2009}, \textsc {LPJ-GUESS}
\citep{Smith-2014} and a\textsc{LPJml-FIT}  \citep{Sakschewski-2015}. These
models were designed for simulation of bio-geochemical cycles and vegetation
structure across broad geographic areas, and their architectures are
optimised for such use. By contrast, {\plant} was  designed so that the
physiological rules driving the dynamics of the system could be easily
manipulated and emergent demographic behaviours (including properties of individual plants, patches , metapopulations of patches and evolutionary fitness) studied in
detail. Finally, like {\plant} the \textsc{LM3-PPA} \cite{Weng-2015} model also aims
to scale from individual-level behaviours to emergent properties of
vegetation, including estimating invasion fitness. Yet the approaches taken
for solving the dynamics are very different. The \textsc{LM3-PPA} relies on
complex analytical solutions to the size- structured interactions, whose
derivation required simplifying elements of the physiology and ecology in the
model. By contrast in {\plant} we have retained the full range of
physiological and ecological feedbacks, and instead relied on using robust
numerical methods to solve the system dynamics (see Appendix
\ref{sec:demography} for further details).

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/plant.pdf}
\caption{Physiological dynamics for individual plants varying in size,
  trait and light environment. Solid lines are for the high
  \textsc{lma} (leaf mass per unit area) strategy, dashed lines are
  for the low \textsc{lma} strategy.  (a) Growth trajectories are
  influenced by both light environment and traits.  (b) Over time, the
  fraction of living tissue switches from leaf towards sapwood, with
  high \textsc{lma} species having relatively more mass in leaf than
  low \textsc{lma} species, due to their longer leaf lifespan. (c) Size-dependent growth rates
  (derivative of panel (a)) peak at around 5m high, but location of the peak
  varies with both trait and light level.  (d) Declines in growth rate
  with light vary with plant size and traits.  Whole plant light
  compensation points (indicated by dots) emerge as the point where
  growth rate is zero (x intercept).  The light level used in panels
  (a) and (c) is indicated with coloured ticks.  See Appendix
  \ref{sec:code} for further details and code.}
\label{fig:plant}
\end{figure}

>>>>

% TODO: Table XXX?
While we have implemented a particular physiological model (\texttt{FFW16}), the
package is designed to allow arbitrary additional physiologies to be
introduced. Users are able to modify the physiological model in two
different ways. First, one can vary the parameters of the \texttt{FFW16} model
directly (see Table XXX for list, for examples see Appendix
\ref{sec:code}). Second, we allow for changes in key parameters to
bring about changes in other parameters (a hyper-parameterisation), enabling
straightforward modelling of trade-offs. For example, the trait \textsc{lma}
(leaf mass per unit leaf area) is used to estimate the rate of leaf turnover
($k_l$), based on an observed scaling relationship spanning across diverse
vegetation and plant functional types \citep{Wright-2004}:
$$k_l = \beta_{kl1} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{kl2}},$$
where $k_l$ and $\phi$ (lma) are both parameters of the core physiological
model, and $\beta_{kl1}, \beta_{kl2}, \phi_0$ are hyper-parameters. Such
linkages are defined within a user-supplied hyper-parameterisation function
(see Appendix \ref{sec:FFW16} for details). Finally, rather than varying the
parameters of the existing equations in the \texttt{FFW16} model, the
equations themselves can be changed and the higher level functionality of the
package can still be used.  Changed equations would amount to defining a new
physiological model in a new C++ file. As an example, in
the \texttt{FFW16b} physiology, we have modified the function determining
allocation to reproduction.

@@

% TODO: Appendix XXX
The {\plant} model offers two methods for modelling the dynamics of \(N(x,
H, a)\) in a competing population. In the first ``stochastic'' mode, users
specify an average seed rain and patch size, and {\plant} then generates a
vector of seed arrival events and simulates stochastic
development of the resulting finite-sized population (see Appendices XXX for an
example).

@@

Increased interest in capturing the behaviour of size-structured  populations
has prompted development of several tools, each with its own distinct
strengths. The \textsc{EBT} software (see De Roos 2016 in this special feature) provides
a flexible framework that has been used to study a range of marine organisms.
Compared to the \textsc{EBT}, {\plant} is more focussed on plants, and
to that end offers refinements needed to handle modelling of the strong
size-asymmetric feedbacks that occur in plant communities, as well as the
extension to structured metapopulation. Other \textsc{TSPMs} also exist for modelling of vegetation, like \textsc{ED}
\citep[ver 1 and 2][]{Moorcroft-2001, Medvigy-2009}, \textsc {LPJ-GUESS}
\citep{Smith-2014} and a\textsc{LPJml-FIT}  \citep{Sakschewski-2015}. These
models were designed for simulation of bio-geochemical cycles and vegetation
structure across broad geographic areas, and their architectures are
optimised for such use. By contrast, {\plant} was  designed so that the
physiological rules driving the dynamics of the system could be easily
manipulated and emergent demographic behaviours (including properties of individual plants, patches , metapopulations of patches and evolutionary fitness) studied in
detail. Finally, like {\plant} the \textsc{LM3-PPA} \cite{Weng-2015} model also aims
to scale from individual-level behaviours to emergent properties of
vegetation, including estimating invasion fitness. Yet the approaches taken
for solving the dynamics are very different. The \textsc{LM3-PPA} relies on
complex analytical solutions to the size- structured interactions, whose
derivation required simplifying elements of the physiology and ecology in the
model. By contrast in {\plant} we have retained the full range of
physiological and ecological feedbacks, and instead relied on using robust
numerical methods to solve the system dynamics (see Appendix
\ref{sec:demography} for further details).

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/plant.pdf}
\caption{Physiological dynamics for individual plants varying in size,
  trait and light environment. Solid lines are for the high
  \textsc{lma} (leaf mass per unit area) strategy (\textsc{lma} =
  0.267), dashed lines are for the low \textsc{lma} strategy
  (\textsc{lma} = 0.08).  (a) Growth trajectories are influenced by
  both light environment and traits.  (b) Over time, the fraction of
  living tissue switches from leaf towards sapwood, with high
  \textsc{lma} species having relatively more mass in leaf than low
  \textsc{lma} species, due to their longer leaf lifespan. (c)
  Size-dependent growth rates (derivative of panel (a)) peak at around
  5m high, but location of the peak varies with both trait and light
  level.  (d) Declines in growth rate with light vary with plant size
  and traits.  Whole plant light compensation points (indicated by
  dots) emerge as the point where growth rate is zero (x intercept).
  The light level used in panels (a) and (c) is indicated with
  coloured ticks.  See Appendix \ref{sec:code} for further details and
  code.}
\label{fig:plant}
\end{figure}
